---
title: "Presentations"
date-modified: today
echo: false
warning: false
message: false
title-block-banner: false
page-layout: article
toc: true
editor_options: 
  chunk_output_type: console
---

# Descriptive Statistics: The Importance of Visualization

* Talk presented at the annual CSHS Biostatistics Fall Lecture

```{r}
knitr::include_url('https://mluu921.github.io/cshs-fall-lecture-descriptive-statistics/#/descriptive-statistics')
```

# Data Manipulation: Introduction to the tidyverse: dplyr & tidyr

* Talk presented for the BMS510 course

```{r}
knitr::include_url('https://mluu921.github.io/cshs-bms510-data-manipulation-slides/#/data-manipulation-in-r')
```

# Descriptive Statistics in R: Introduction to plotting in the tidyverse: ggplot2

* Talk presented for the BMS510 course

```{r}
knitr::include_url('https://mluu921.github.io/cshs-bms510-ggplot2-slides/#/descriptive-statistics-in-r')
```

# Descriptive Statistics

* Talk presented for the MRM615 course

```{r}
knitr::include_url('https://mluu921.github.io/biostats_mrm615_descriptive/')
```

