---
title: "Retrieve Pubmed publications using the RefManageR package"
date: "4/13/2023"
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
library(tidyverse)
library(RefManageR)
library(ggwordcloud)

```

For anyone that's working in academia, it may be useful to keep a tabulation of all of the publications that you are a co-author of. We can use the `RefManageR` package, which includes a function (`ReadPubMed()`) that allows us to query the PubMed API for publications based on a PubMed query.

Below I generate a PubMed query `q` that would query PubMed for publications that involves `Luu, Michael` as an author with a `Cedars-Sinai` affiliation. I then pass this query into the `ReadPubMed()` function and save the results as `pm`. The output is a `BibEntry` object that can be further coerced into a tibble.

```{r}

q <- '(Luu, Michael[Author]) AND (Cedars-Sinai[Affiliation])'

pm <- RefManageR::ReadPubMed(q, retmax = 999)
  
out <- pm |> as_tibble()

glimpse(out)

out
```

Now that we have this information into a tibble, we can further visualize the frequency of occurrences among the list of co-authors using a word cloud

```{r}
#| out-width: 100%

plot_data <- out |>
  select(author) |>
  separate_wider_delim(
    author,
    names = paste0('author', 1:20),
    delim = ' and ',
    too_few = 'align_start'
  ) |>
  mutate(i = row_number()) |>
  pivot_longer(contains('author')) |>
  filter(!is.na(value)) |> 
  count(value) |> 
  arrange(desc(n)) |> 
  filter(value != 'Michael Luu')

set.seed(1)
ggplot(plot_data, aes(label = value, size = n, color = n)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 16) +
  theme_minimal() +
  scale_color_viridis_c()

```

# Session info

```{r}
sessionInfo()
```
