---
title: "Haunted Los Angeles - Interactive Mapping using Leaflet"
date: '11/21/2023'
warning: false
message: false
image: "https://images.unsplash.com/photo-1481018085669-2bc6e4f00eed?q=80&w=3072&auto=format&fit=crop&ixlib=rb-4.0.3&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D"
categories: 
  - tidytuesday
  - leaflet
  - gis
  - visualization
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(tidytuesdayR)
library(leaflet)

```

The following blog post is a very simple visualization using the data from tidytuesday 2023, week 41. This visualization uses the leaflet library to produce a very simple spatial analysis of the 'haunted' locations found in Los Angeles, CA.

We begin by loading the data from the tidytuesdayR package.

```{r}

datas <- tt_load(2023, 41)

data <- datas$haunted_places

```

Next we take a glimpse of the data to see what we are working with.

```{r}
glimpse(data)
```

We next subset the data for the observations found in Los Angeles, CA. The longitude and latitude are then used to plot the locations on a map using the leaflet library. The 'J.F.K. Library Third' location is removed from the data as it is an outlier and is not located in Los Angeles, CA.

```{r}
#| column: page
#| fig-height: 10

data |>
  filter(city %in% c('Los Angeles')) |>
  filter(state_abbrev == 'CA') |>
  filter(location != 'J.F.K. Library Third') |> 
  mutate(label = glue::glue('<b>{location}</b> <br> {description}')) |> 
  leaflet() |>
  addTiles() |>
  addMarkers(
    lng = ~ longitude,
    lat = ~ latitude,
    popup = ~ label
  )

```
